A Case Study on Abstractions to Converge High-Throughput and High-Performance Computing

### 2017-03-19 bckground.tex

The Production and Distributed Analysis (PanDA) is a Workload Management System
(WMS) developed to support the operations of the LHC ATLAS experiment. PanDA is
specifically design to manage the execution of distributed workloads and
workflows via pilots.

WMS for distributed executions are a type of middlware, designed to coordinate
specific activities about resources and workloads: discovering and selecting
resources, submitting the tasks of a workload, and monitoring the execution of
those tasks~\cite{marco2009glite}. Pilot is an abstraction that enables
multi-level scheduling by decoupling resource acquisition from workload
scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
placeholders: a request for a certain amount of resources is scheduled on a site
(e.g., by submitting a job or requesting a virtual machine) and once the
resources are acquired, tasks are scheduled directly to the pilot, not to site's
scheduler.

WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
resources by means of pilots and then manage the execution of workloads or
workflows' tasks on those pilots. In tis way, pilots expose a uniform scheduling
interface and execution environment, abstracting the differences among the
heterogeneous resources of multiple sites; WMS centralize the management of the
execution process, offering a unified interface to the application layer.
Pilot-enabled WMS are particularly relevant for LHC experiments, where large
amount of heterogeneous resources, distributed across multiple sites have to be
coordinated to execute tens of million of tasks per months while producing and
storing petabytes of data.

Several pilot-enabled WMS were developed for the LHC experiments:
AliEn~\cite{Bagnasco2010} for ALICE; DIRAC~\cite{Paterson2010} for LHCb;
GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS; and
PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
design and architectural principles: separation of concern between management
and execution, and data and compute; centralization of the management
capabilities; distributed execution of the execution capabilities across
multiple sites; unification of the application interface; hiding of resource
heterogeneity; centralized monitoring and accounting of both resources and
execution; and collection of static and sometimes dynamic information about
resources.

AliEn, DIRAC, GlideinWMS and PanDA all share a similar design with two types of
components: the management ones facing the application layer and centralizing
the capabilities required to acquire tasks' descriptions and matching them to
resource capabilities; and resource components required to acquire information
about resource capabilities and the resources themselves. Architecturally, the
management components include one or more queue and a scheduler capable of
coordinating via push/pull protocols with the resource modules. All resource
components include middleware-specific APIs for submitting resource requests
and, in the case of the pilot-enabled WMS of the LHC projects, a pilot capable
of pulling tasks from the management modules and scheduling them on their
resources (i.e., multi-level scheduling).

AliEn, DIRAC, GlideinWMS and PanDA present several similarities also in their
implementations. All were initially implemented to use Grid resources, using one
or more components to the Condor software ecosystem~\cite{thain2005distributed}
and, as with GlideinWMS, contributing to its development. Accordingly, all LHC
WMS implemented Grid-like authentication and authorization systems and adopted a
computational model based on distributing a large amount of single/few-cores
tasks across hundreds of sites\mtnote{Is this true?}. Tasks are assumed to fail
or being held and all WMS implements subsystems dedicated to manage these
failures or simply resubmitting failed tasks.

@@

Currently, the implementation of PanDA WMS has at least two distinguishing
features: the scale at which it manages computation, and network-aware
scheduling of tasks to sites. PanDA concurrently supports distributed
computations on up to 240,000 cores, submitting more than 365 million jobs in
2016 across more than 100 sites, and processing up to XXXXPB of data distributed
worldwide across XX Tier 1-3 facilities. As for 2017, PanDA serves several
thousand users, including XX production groups. The scale of this operation is
unprecedented, making PanDA the WMS managing the largest computational campaign
in the world.

@@

The dynamic management of data replication and download is an example of a
generalized paradigm shift that distinguishes the implementation of PanDA from
other WMS. New components are being implemented to support the dynamic sizing of
input dataset based on resource capabilities, the dynamic sizing of the
resources held by pilots both in terms of number of cores and duration of the
pilot, and further abstraction of different types of resources to reconcile
High-Throughput and High-Performance Computing paradigms.

@@

% The terms ``task'' and ``job'' are also used inconsistently across communities
% that perform scientific computing. In this paper, ``job'' refers to the unit of
% work that is submitted to a local resource management system (LRMS), like the
% Slurm or PBS batch system of a cluster. As such, a task can become a job when
% is scheduled on a resource that exposes a LRMS but can also become a virtual
% machine or a container when bootstrapped on an infrastructure supporting
% virtualization. Tasks can be statically or dynamically grouped into jobs,
% depending on the resource capabilities and the task requirements.
%
% Usually, workflows are represented as graphs in which tasks are vertices and
% relations are edges~\cite{}. Often, graphs are supposed to be acyclic but graphs
% with cycles have been used to represent workflows of workflows~\cite{}. In this
% paper, a workflow template is a type of graph while a workflow instance is a
% workflow template with specific vertices and edges. Further, a workflow instance
% is ``abstract'' when no resource properties are available for all vertices,
% ``concrete'' otherwise~\cite{}. When not qualified, the term ``workflow''
% indicates a abstract workflow instance.
%
% Notes for refactoring:
% \begin{itemize}
%     \item Naming and nomenclature is a contribution
%     \item 1, 1.5 pages.
% \end{itemize}

>>>>

The Production and Distributed Analysis (PanDA) software is a Workload
Management System (WMS) developed to support the operations of the LHC ATLAS
experiment. PanDA is specifically design to manage the execution of distributed
workloads and workflows via pilots.

WMS for distributed executions are a type of middleware, designed to coordinate
specific activities about resources and workloads: discovering and selecting
resources, submitting tasks of workloads and workflows, and monitoring the
execution of those tasks~\cite{marco2009glite}. Pilot is an abstraction that
enables multi-level scheduling by decoupling resource acquisition from workload
scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
placeholders: a request for a certain amount of resources is scheduled on a site
(e.g., by submitting a job or requesting a virtual machine) and once the
resources are acquired, tasks are scheduled directly to the pilot, not to site's
scheduler.

WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
resources by means of pilots and then manage the execution of workloads or
workflows' tasks on those pilots. WMS centralize the management of the execution
process, offering a unified interface to the application layer. Pilots expose a
uniform scheduling interface and execution environment, abstracting the
differences among the heterogeneous resources of multiple sites. Pilot-enabled
WMS are particularly relevant for LHC experiments, where millions of tasks are
executed across multiple sites every month, analyzing and producing petabytes of
data.

Several pilot-enabled WMS were developed for the LHC experiments:
AliEn~\cite{Bagnasco2010} for ALICE; DIRAC~\cite{Paterson2010} for LHCb;
GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS; and
PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
design and architectural principles: centralization of task and resource
management, and of monitoring and accounting; distribution of task execution
across multiple sites; unification of the application interface; hiding of
resource heterogeneity; and collection of static and sometimes dynamic
information about resources.

AliEn, DIRAC, GlideinWMS and PanDA all share a similar design with two types of
components: the management ones facing the application layer and centralizing
the capabilities required to acquire tasks' descriptions and matching them to
resource capabilities; and resource components used to acquire compute and data
resources and information about their capabilities. Architecturally, the
management components include one or more queue and a scheduler that coordinates
with the resource modules via push/pull protocols. All resource components
include middleware-specific APIs to request for resources, and a pilot capable
of pulling tasks from the management modules and executing them on its
resources.

AliEn, DIRAC, GlideinWMS and PanDA also have similar implementations. These WMS
were initially implemented to use Grid resources, using one or more components
to the Condor software ecosystem~\cite{thain2005distributed} and, as with
GlideinWMS, contributing to its development. Accordingly, all LHC WMS
implemented Grid-like authentication and authorization systems and adopted a
computational model based on distributing a large amount of single/few-cores
tasks across hundreds of sites\mtnote{Is this true?}.

@@

Currently, the implementation of PanDA WMS has at least two distinguishing
features: the scale at which it operates distributed computing, and
network-aware scheduling of tasks to sites\mtnote{PanDA team: please let me know
whether more unique (i.e., not present in other LHC WMS) features of PanDA
implementation should be indicated}. PanDA concurrently supports distributed
computations on up to 240,000 cores, submitting more than 365 million jobs a
year across 100 sites, and processing up to XXXXPB of data distributed worldwide
on XX Tier 1-3 facilities. As for 2017, PanDA serves several thousand users,
including XX production groups. The scale of this operation is unprecedented,
making PanDA the WMS that manages the largest computational campaign in the
world.

@@

The dynamic management of data replication and download is an example of a
generalized paradigm shift that distinguishes the implementation of PanDA from
other WMS. New components are being implemented to support: (i) dynamic sizing
of input dataset based on resource capabilities; (ii) dynamic sizing of the
resources held by pilots both in terms of number of cores and duration of the
pilot; and (iii) further abstraction of different types of resources to
reconcile high-throughput and high-performance computing paradigms.

@@

% The terms ``task'' and ``job'' are also used inconsistently across communities
% that perform scientific computing. In this paper, ``job'' refers to the unit
% of work that is submitted to a local resource management system (LRMS), like
% the Slurm or PBS batch system of a cluster. As such, a task can become a job
% when is scheduled on a resource that exposes a LRMS but can also become a
% virtual machine or a container when bootstrapped on an infrastructure
% supporting virtualization. Tasks can be statically or dynamically grouped
% into jobs, depending on the resource capabilities and the task requirements.

% Usually, workflows are represented as graphs in which tasks are vertices and
% relations are edges~\cite{}. Often, graphs are supposed to be acyclic but
% graphs with cycles have been used to represent workflows of
% workflows~\cite{}. In this paper, a workflow template is a type of graph
% while a workflow instance is a workflow template with specific vertices and
% edges. Further, a workflow instance is ``abstract'' when no resource
% properties are available for all vertices, ``concrete'' otherwise~\cite{}.
% When not qualified, the term ``workflow'' indicates a abstract workflow
% instance.

% Notes for refactoring:
% \begin{itemize}
%     \item Naming and nomenclature is a contribution
%     \item 1, 1.5 pages.
% \end{itemize}

### 2017-03-19 nge.tex

As seen in \S\ref{sec:panda_deployment}, the absence of wide area network
connectivity on Titan's work nodes was one of the main reasons for PanDA Broker
not to implement pilots. As discussed in \S\ref{sec:panda_titan}, the absence of
pilots imposes the static coupling between PBS jobs and detector simulations,
and development tailored to each type of payload that needs to be executed.

@@

Pilots can offer a payload-independent scheduling interface while hiding the
mechanics of coordination and communication among multiple work nodes. This
could eliminate the need for packaging payload into PBS jobs within the broker,
greatly simplifying the submission process. This simplification would also
enable the submission of different types of payload, without having to develop a
specific PBS script for each payload. The submission process would also be
MPI-independent, as MPI is used for coordination among multiple work nodes, not
by the payload.

>>>>

As seen in \S\ref{sec:panda_deployment}, the absence of wide area network
connectivity on Titan's worker nodes was one of the main reasons for PanDA
Broker not to implement pilots. As discussed in \S\ref{sec:panda_titan}, the
absence of pilots imposes the static coupling between PBS jobs and detector
simulations, and development tailored to each type of payload that needs to be
executed.

@@

Pilots can offer a payload-independent scheduling interface while hiding the
mechanics of coordination and communication among multiple worker nodes. This
could eliminate the need for packaging payload into PBS jobs within the broker,
greatly simplifying the submission process. This simplification would also
enable the submission of different types of payload, without having to develop a
specific PBS script for each payload. The submission process would also be
MPI-independent, as MPI is used for coordination among multiple worker nodes,
not by the payload.

### 2017-03-19 overall.tex

Usually, detector simulations performed on the Grid process 10 times the number
of events than those on Titan. This explains the difference between the
percentage of detector simulations (9\%) and of events computed by those
simulations (3.9\%) on Titan. PanDA Broker could fit the number of events to the
walltime of the backfill availability on the base of the distribution of
Figure~\ref{fig:backfill-distrib}. That specific number of event could then be
pulled from the PanDA Event service~\cite{panda-event-service} and given as
input to one or more simulations. Once packaged into the MPI script submitted to
titan's PBS batch system, these simulations would better fit backfill
availability, increasing the efficiency of PanDA Brokers.

>>>>

Usually, detector simulations performed on the Grid process 10 times the number
of events than those on Titan. This explains the difference between the
percentage of detector simulations (9\%) and of events computed by those
simulations (3.9\%) on Titan. PanDA Broker could fit the number of events to the
walltime of the backfill availability on the base of the distribution of
Figure~\ref{fig:backfill-distrib}. That specific number of event could then be
pulled from the PanDA Event service~\cite{calafiura2015atlas} and given as input
to one or more simulations. Once packaged into the MPI script submitted to
titan's PBS batch system, these simulations would better fit backfill
availability, increasing the efficiency of PanDA Brokers.

### 2017-03-19 paper.tex
Ongoing section II
proposed new title.
First pass section II; moved definitions to end of introduction

\documentclass[10pt, conference, compsocconf]{IEEEtran}

@@

@@

\title{Everything you wanted to know about PanDA but were truly truly afraid to
ask}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations
\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
\and
\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

@@

\begin{IEEEkeywords}
TBD\@.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

% ----------------------------------------------------------------------------
% II - BACKGROUND AND RELATED WORK
% ----------------------------------------------------------------------------
\section{Background and Related Work}
\label{sec:bkgrd}

\input{WMSLiterature}
% Please do not correct bckground to background.
\input{bckground}

@@

% ----------------------------------------------------------------------------
% REFERENCES
% ----------------------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{bibliography}

>>>>

\documentclass[sigconf, screen]{acmart}

@@

% DOI
% \acmDOI{10.475/123_4}

% ISBN
% \acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[SC2017]{The International Conference for High Performance Computing, Networking, Storage and Analysis}{November 2017}{Denver, Colorado USA}
\acmYear{2017}
\copyrightyear{2017}

% \acmPrice{15.00}

@@

\title{A Case Study on Abstractions to Converge High-Throughput and High-Performance Computing}

\author{Author 1}
\affiliation{\institution{Institution 1}}
\email{email1}

\author{Author 2}
\affiliation{\institution{Institution 2}}
\email{email2}

\author{Author 3}
\affiliation{\institution{Institution 3}}
\email{email3}

\author{Author 4}
\affiliation{\institution{Institution 4}}
\email{email4}

\author{Author 5}
\affiliation{\institution{Institution 5}}
\email{email5}

\author{Author 6}
\affiliation{\institution{Institution 6}}
\email{email6}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{Author 1 et al.}

@@

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
\end{CCSXML}

% \ccsdesc[500]{Computer systems organization~Embedded systems}
% \ccsdesc[300]{Computer systems organization~Redundancy}
% \ccsdesc{Computer systems organization~Robotics}
% \ccsdesc[100]{Networks~Network reliability}

\keywords{ACM proceedings, \LaTeX, text tagging}

\maketitle

@@

% ----------------------------------------------------------------------------
% II - RELATED WORK
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}

% \input{WMSLiterature}
% Please do not correct bckground to background.
\input{bckground}

@@

% ----------------------------------------------------------------------------
% REFERENCES
% ----------------------------------------------------------------------------
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}
